---
title: "SPLANG Quickstart Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SPLANG Quickstart Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

SPLANG is a **s**ynthetic **P**oisson-**l**ognormal-based **a**bundance and 
**n**etwork **g**enerative model. SPLANG extends the work from 
(Chiquet et.al 2019) by enabling the expression of asymmetric taxa interaction
and can generate synthetic abundance sampling data along with the underlying
taxa interaction network. SPLANG generates pairs of abundance data and underlying
taxa interaction network can be used to benchmark and inform the research
and development of taxa inference algorithms.

```{r setup}
library(SPLANG)
```

## Generate Paired Synthetic Abundance Data and the Underlying Taxa Interaction Network

```{r simple-example}
sample.sequencing(
  n.species=5,
  n.sequencing=10,
  sequencing.depth=100,
  network.type = 'nm',
  m.edges=(5 * (5 - 1) * .5) # use the floor value if not an integer
)
```


## Generate in Batch

To generate in batches, we can create a helper function like

```{r an-example-of-a-helper-function}
library(stringr)

generate.synthetic.data = function(
    n.species,
    n.sequencing,
    sequencing.depth,
    n.sample.files,
    dir.path,
    fn.opath = function(n.species, n.files, i) {
      file.path(
        dir.path,
        str_glue("apc-{n.species}-species-{i}-of-{n.sample.files}.Rdata")
      )
    },
    ...
) {
  if (!dir.exists(dir.path)) {
    dir.create(dir.path, recursive = T)
  }
  for (i in 1:n.sequencing) {    
    gen.data = sample.sequencing(n.species, n.sequencing, sequencing.depth, ...)
    save(gen.data, file=fn.opath(n.species, n.sample.files, i))
  }
}
```

The following code will generate batches of abundance data and the underlying
taxa interaction network for 12, 25, 50, 100, and 200 species, each with 100
realizations of abundance data and underlying taxa interaction network.

The generated file will be stored under a folder named `gen-data` under the
current working directory, grouped by the number of species as subdirectories.

For each realization, the abundance data and the underlying taxa interaction network
will be stored in `.Rdata` and can be loaded with `load(path_to_Rdata_file)`.

```{r generate-in-batch}
all.num.of.species = c(12, 25, 50, 100, 200)

## WARNING: The following code is commented because it will generate 500 files
## under the current working directory. Please run with caution.

# for (num.species in all.num.of.species) {
# 
#   generate.synthetic.data(
#     n.species = num.species,
#     n.sequencing = 100,
#     sequencing.depth = 30000,
#     n.sample.files = 100,
#     dir.path = str_glue("gen-data/data-apc-pa-{num.species}-species-30k-depth")
#   )
#     
# }
```